---
title: "ICRW8 Workshop: Integrating USDA-ARS and NEON Water Quality Data for Southeast Rivers"
description: "Tutorial for downloading NEON data and USDA ARS data, and then exploring and understanding the downloaded data"
dateCreated: 2023-05-08
authors: Bobby Hensley, Jim Coloso
contributors: 
estimatedTime: 120 minutes
packagesLibraries: neonUtilities, plotly, lubridate
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
---
```{css toc-content, echo = FALSE}
/*set set margins so that TOC all the way to the left*/
#TOC {
  /*right: 270px;*/
  margin: 20px 0px 25px 0px;
}
.main-container {
    margin-left: 20px;
}
body .main-container{
    max-width: 90vw;
}
```
```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = T, echo = T, message = F, warning = F, error = F)
```

This tutorial covers downloading NEON Aquatic Instrument System (AIS)
and Aquatic Observation System (AOS) data from the Flint River site
(FLNT) as well as Little River data from the USDA-ARS program. This also
covers data analysis, plotting, and a comparison of water quality
parameters at the two sites.

::: {#ds-objectives markdown="1"}
## Objectives

After completing this activity, you will be able to:

-   Download NEON AIS and AOS data using the `neonUtilities` package.
-   Understand downloaded data sets and load them into R for analyses.
-   Merge instrumented an observational data sets for plotting and
    comparison.
-   Create concentration-discharge plots    

## Things You'll Need To Complete This Tutorial

To complete this tutorial you will need R (version \>=4.2) and,
preferably, RStudio loaded on your computer.

### Install R Packages

-   **neonUtilities**: Basic functions for accessing NEON data
-   **plotly**: Plotting functions
-   **lubridate**: Functions for handling dates and times

These packages are on CRAN and can be installed by `install.packages()`.

### Additional Resources

-   <a href="https://github.com/NEONScience/NEON-Utilities/neonUtilities" target="_blank">GitHub
    repository for neonUtilities</a>
:::

## NEON Flint River Data 

### Download and Load Directly to R

The most popular function in `neonUtilities` is `loadByProduct()`. This
function downloads data from the NEON API, merges the site-by-month
files, and loads the resulting data tables into the R environment,
assigning each data type to the appropriate R class. This is a popular
choice because it ensures you're always working with the most up-to-date
data, and it ends with ready-to-use tables in R. However, if you use it
in a workflow you run repeatedly, keep in mind it will re-download the
data every time.

Before we get the NEON data, we need to install (if not already done)
and load the neonUtilities R package, as well as other packages we will
use in the analysis.

```{r set-up-env, eval=FALSE}
# Install neonUtilities package if you have not yet.
install.packages('neonUtilities')
install.packages('plotly')
install.packages('luridate')
```

```{r load-packages}
# Load required packages
library(neonUtilities)
library(plotly)
library(lubridate)
```

The inputs to `loadByProduct()` control which data to download and how
to manage the processing. The following are frequently used inputs:

-   `dpID`: the data product ID, e.g. DP1.20288.001. It is the data product identifier of the data you want to
    download. It will be in the form DP#.#####.###. For this tutorial, we'll use some data products collected
    in NEON's Aquatic Instrument and Observational Systems: DP1.20033.001 - SUNA nitrate sensor data, DP1.20093.001 - Chemical
    properties of surface water, DP1.20288.001 - Water quality sensor data, and DP4.00130.001 - Continuous discharge.

-   `site`: defaults to "all", meaning all sites with available data; To
    download data from a specific site, use the 4-letter NEON site code.
    The site code for the Flint River is FLNT.

-   `startdate` and `enddate`: defaults to NA, meaning all dates with
    available data; or a date in the form YYYY-MM, e.g. 2022-06. Since
    NEON data are provided in month packages, finer scale querying is
    not available. Both start and end date are inclusive.

-   `package`: either basic or expanded data package. Expanded data
    packages generally include additional information about data
    quality, such as individual quality flag test results. Not every
    NEON data product has an expanded package; if the expanded package
    is requested but there isn't one, the basic package will be
    downloaded.

-   `check.size`: TRUE or FALSE; should the function pause before
    downloading data and warn you about the size of your download?
    Defaults to TRUE; if you are using this function within a script or
    batch process you will want to set this to FALSE. For large or
    unknown downloads, you may want to check the size first.

-   `token`: this allows you to input your NEON API token to obtain
    faster downloads. This is is optional and will not be used below for
    simplicity.

Learn more about NEON API tokens in the
<a href="https//:www.neonscience.org/neon-api-tokens-tutorial" target="_blank">**Using
an API Token when Accessing NEON Data with neonUtilities** tutorial</a>.

There are additional inputs you can learn about in the
<a href="https//:www.neonscience.org/neonDataStackR" target="_blank">**Use
the neonUtilities R Package to Access NEON Data** tutorial</a>.

Now let us download our data. In this exercise, we want data from only one NEON field site, Flint
River (FLNT) from 2018. If you are using a NEON token to download
your data, paste it in the line below and uncomment.

```{r download-data-nsw, results='hide', cache=TRUE}
# Download SUNA nitrate sensor data
nsw <- neonUtilities::loadByProduct(dpID = "DP1.20033.001",
                                    site = "FLNT",
                                    startdate = "2018-01",
                                    enddate = "2018-12",
                                    package = "expanded",
                                    # token = 'paste your token here',
                                    check.size = F
                                    )
```

Using what you've learned above, can you modify the code to download
data for: DP1.20288.001 - Water quality sensor data, DP4.00130.001 Continuous discharge, and  DP1.20093.001 - Chemical
properties of surface water

```{r download-data-others, results='hide', cache=TRUE}
# Download water quality sensor data
waq <- neonUtilities::loadByProduct(dpID = "DP1.20288.001",
                                     site = "FLNT",
                                     startdate = "2018-01",
                                     enddate = "2018-12",
                                     package = "expanded",
                                     check.size = F)

# Download continuous discharge data 
csd <- neonUtilities::loadByProduct(dpID = "DP4.00130.001",
                                    site = "FLNT",
                                    startdate = "2018-01",
                                    enddate = "2018-12",
                                    package = "basic", #try "expanded" to get additional data
                                    check.size = F)

# Download water chemistry data
swc <- neonUtilities::loadByProduct(dpID = "DP1.20093.001",
                                    site = "FLNT",
                                    startdate = "2018-01",
                                    enddate = "2018-12",
                                    package = "expanded",
                                    check.size = F)
```

#### Files Associated with Downloads

The data we've downloaded comes as an object that is a named list of
objects. We can use the `names()` function to view the components of each list.
You can see that there are 8 objects: 1 dataframe of data (`NSW_15_minute`) and
6 metadata files. We can then use the `$` operator to select a specific object from the list. 

```{r loadBy-list nsw}
# view all components of the list
names(nsw)

# View the dataFrame
View(nsw$NSW_15_minute)
```

Now let's view the objects in the other lists.

```{r loadBy-list all}
names(waq)
names(csd)
names(swc)
```

If you'd like, you can use the `$` operator to assign an object from an
item in the list. If you prefer to extract each table from the list and
work with it as independent objects, which we will do, you can use the
`list2env()` function.

```{r unlist-vars,results='hide'}
# Unlist the variables and add to the global environment
list2env(nsw, .GlobalEnv)
list2env(waq, .GlobalEnv)
list2env(csd, .GlobalEnv)
list2env(swc, .GlobalEnv)
```

So what exactly are these files and why would you want to use them?

-   **data file(s)**: There will always be one or more dataframes that
    include the primary data of the data product you downloaded.
    Multiple dataframes are available when there are related datatables
    for a single data product.
-   **readme_xxxxx**: The readme file, with the corresponding 5 digits
    from the data product number, provides you with important
    information relevant to the data product.
-   **sensor_postions_xxxxx**: this file contains information about the
    coordinates of each sensor, relative to a reference location.
-   **variables_xxxxx**: this file contains all the variables found in
    the associated data table(s). This includes full definitions, units,
    and other important information.
-   **issueLog_xxxxx**: this file contains information about issues
    identified for the particular data product.

There are also two files specific to water quality:

-   **ais_maintenance**: this file contains information about field
    maintenance, for example when cleanings and calibrations were
    performed.
-   **ais_multisondeCleanCal**: this file contains information from the
    cleanings and claibrations, for example what the sensor was reading
    in standards before and after. This can be useful in performing
    drift correction or adjusting calibration offsets.

The water chemistry data also contains:

-   **validation_xxxxx**: this file provides descriptions of data validation.

#### A note about sensor positions

NEON often collects the same type of data from sensors in different
locations. These data are delivered together but you will frequently
want to plot the data separately or only include data from one sensor in
your analysis. NEON uses the `horizontalPosition` variable in the data
tables to describe which sensor data is collected from. The
`horizontalPosition` is always a three digit number for AIS data.
Examples as of 2022 at AIS sites include:

-   101: stream sensors located at the **upstream** station on a
    **monopod mount**,
-   111: stream sensors located at the **upstream** station on an
    **overhead cable mount**,
-   131: stream sensors located at the **upstream** station on a **stand
    alone pressure transducer mount**,
-   102: stream sensors located at the **downstream** station on a
    monopod mount,
-   112: stream sensors located at the **downstream** station on an
    **overhead cable mount**
-   132: stream sensors located at the **downstream** station on a
    **stand alone pressure transducer mount**,
-   110: **pressure transducers** mounted to a **staff gauge**.
-   103: sensors mounted on **buoys in lakes or rivers**
-   130 and 140: sensors mounted in the **littoral zone** of lakes

The Flint River data in this exercise is collected from a single buoy so
it all has the same `horizontalPosition` of 103. `horizontalPosition` is
much more important when looking at stream or groundwater data where
there are multiple locations.

### Compare nitrate sensor data to nitrate from grab samples

#### Examine and prep nitrate sensor data

First, let's identify the column names important for our analysis - time
and nitrate data. There are several ways of doing this:

```{r column-names}
# One option is to view column names in the data frame
names(NSW_15_minute)

# Alternatively, view the variables object corresponding to the data product for more information
View(variables_20033)
```

The time column we'll consider for instrumented systems is `endDateTime`
because it approximately represents data within the interval on or
before the `endDateTime` time stamp. Timestamp column choice matters for
time-aggregated datasets, but should not matter for instantaneous data
such as water quality. When interpreting data, keep in mind NEON
timestamps are always in UTC.

The data column we would like to plot is labeled `surfWaterNitrateMean`.

First, we need to remove any quality flags in the nitrate data. There
are many different quality flags in NEON sensor data, but they are
summarized in the finalQF column where 0 has no flag and 1 has a flag.
We will remove any `finalQF` flags where the value is not 0.

```{r nsw remove flags}
# Remove flagged data
NSW_15_minute <- NSW_15_minute[NSW_15_minute$finalQF == 0,]
```


Now let's create a plot. We're using the `plotly` package as it allows
for interactive data exploration. We can also add in the uncertainty
values. In the plot below, you can use the tools in the upper right-hand
corner to zoom, pan, etc. You can turn traces on/off by clicking on them
in the legend. You can also see information about the data by hovering the 
cursor over the plot.

```{r plot nsw}
plot_ly(data = NSW_15_minute, 
        x = ~endDateTime,
        y = ~surfWaterNitrateMean,
        name = "SUNA",
        type = "scatter",
        mode = "lines") %>%
  add_ribbons(data = NSW_15_minute, 
              x = ~endDateTime,
              ymin = ~(surfWaterNitrateMean - surfWaterNitrateExpUncert), 
              ymax = ~(surfWaterNitrateMean + surfWaterNitrateExpUncert),
              name = "Uncertainty"
              ) %>%
  layout(title = 'SUNA Nitrate with Uncertainty',
         xaxis = list(title = "Date"),
         yaxis = list(title = "Nitrate (micromoles/L)")
         )

```

Next, we will need to convert nitrate from µM to nitrogen mg/L so we can
compare with grab samples.

```{r nsw convert N}
# Convert to mg/L
NSW_15_minute$nitrate_mgL = NSW_15_minute$surfWaterNitrateMean*14/1000
```


Now let's get the grab sample nitrate data ready for comparison. We need
to pull out the nitrate data and match it up with the SUNA data.

```{r swc nitrate}
# Get nitrate (NO3+NO2 - N)
swc_nitrate<-swc_externalLabDataByAnalyte[swc_externalLabDataByAnalyte$analyte=="NO3+NO2 - N",]

# Average by date (some replicates)
swc_nitrate <- setNames(aggregate(swc_nitrate$analyteConcentration, 
                                  by = list(swc_nitrate$collectDate), 
                                  FUN = mean), 
                        c('collectDate','analyteConcentration'))

# Round date to nearest 15 min to match SUNA
swc_nitrate$roundedDate<-lubridate::round_date(swc_nitrate$collectDate,unit="15 minute")


```

We can now plot SUNA nitrate with grab sample nitrate

```{r suna swc plot}
# Create plot of SUNA and grab sample time-series
plot_ly(data = NSW_15_minute, 
        x = ~endDateTime,
        y = ~nitrate_mgL,
        name = "SUNA",
        type = "scatter",
        mode = "lines") %>%
  add_trace(data = swc_nitrate, 
            x = ~collectDate,
            y = ~analyteConcentration,
            name="Grab Sample",
            type="scatter",
            mode="markers") %>%
  layout(title = 'SUNA Nitrate with Grab Samples',
         xaxis = list(title = "Date"),
         yaxis = list(title = "NO2+NO3-N (mg/L)"))
         

```

The two data sets seem to correspond fairly well. Let's merge the data 
and run a regression to compare them.

```{r nitrate reg}
# Merge with SUNA with grab samples
mergedData<-merge(swc_nitrate,
                  NSW_15_minute,
                  by.x="roundedDate",
                  by.y = 'endDateTime',
                  all = TRUE
                 )
# Create regression of SUNA vs grab samples and fits linear model
sunaGrabReg<-na.omit(mergedData[,c("analyteConcentration","nitrate_mgL")]) #only matching pairs
fit<-lm(nitrate_mgL ~ analyteConcentration, data = sunaGrabReg)
plot_ly(data = sunaGrabReg, 
        x = ~analyteConcentration,
        y = ~nitrate_mgL,
        name = "data",
        type = "scatter",
        mode = "markers")%>%
  add_trace(data=sunaGrabReg,
            x = ~analyteConcentration,
            y = fitted(fit),
            name = "regression",
            mode = "lines")%>%
  layout(title = 'SUNA vs Grab Samples',
         xaxis = list(title = "grab NO2+NO3-N (mg/L)"),
         yaxis = list(title = "SUNA NO2+NO3-N (mg/L)")
         )
summary(fit)
```

### Compare fDOM sensor data to DOC from grab samples

Now, let's follow the same steps to compare DOC from grab samples with
in-situ fDOM measurements.

```{r doc fdom}
# Get fDOM from water quality sensor data and remove flags
fdom <- waq_instantaneous[waq_instantaneous$fDOMFinalQF == 0,]
fdom <- fdom[,c("endDateTime", "fDOM", "rawCalibratedfDOM")]

# Get DOC from water chemistry data
swc_doc <- swc_externalLabDataByAnalyte[swc_externalLabDataByAnalyte$analyte == "DOC",]

# Average by date (some replicates)
swc_doc <- setNames(aggregate(swc_doc$analyteConcentration, 
                              by = list(swc_doc$collectDate), 
                              FUN = mean), 
                    c('collectDate','analyteConcentration'))


# Plot fDOM and grab samples
## Create a second Y-axis to make plot easier to read
y2 <- list(
  overlaying = 'y',
  side =  'right',
  title = 'DOC mg/L'
)

## Plot fDOM with DOC
plot_ly(data = fdom, 
        x = ~endDateTime,
        y = ~fDOM,
        name="meanfDOM",
        type = "scatter",
        mode = "lines") %>%
  add_trace(data = swc_doc, 
            x = ~collectDate,
            y = ~analyteConcentration,
            name = "grabSample",
            type = "scatter",
            mode = "markers",
            yaxis = 'y2') %>%
  layout(title = 'fDOM and DOC',
         yaxis2 = y2,
         xaxis = list(title = "Date"),
         yaxis = list(title = "fDOM")
  )

# Merge data in order to run regression
## fDOM measured every 1 minute, but lots of flags so not all minutes present - calculate 5 minute averages
fdom$roundedDate <- lubridate::round_date(fdom$endDateTime, unit = '5 minute')
fdom <- setNames(aggregate(fdom$fDOM, 
                           by = list(fdom$roundedDate), 
                           FUN = mean), 
                 c('roundedDate', 'meanfDOM'))

## Round date to nearest 5 min to match fDOM
swc_doc$roundedDate<-lubridate::round_date(swc_doc$collectDate,unit = "5 minute")


## Merge fDOM with DOC data
mergedData.doc<-merge(swc_doc,
                      fdom,
                      by = "roundedDate",
                      all.x = TRUE
                      )


# Plot DOC vs fdom and run regression
docfDOMReg <- na.omit(mergedData.doc[,c('analyteConcentration', 'meanfDOM')])
fit.doc.fdom <- lm(analyteConcentration ~ meanfDOM, data = docfDOMReg)
plot_ly(data = docfDOMReg, 
        x = ~meanfDOM,
        y = ~analyteConcentration,
        name = "data",
        type = "scatter",
        mode = "markers")%>%
  add_trace(data = docfDOMReg, 
            x = ~meanfDOM,
            y = ~predict(fit.doc.fdom),
            name = "regression",
            type = "scatter",
            mode = "lines") %>%
    layout(title = 'fDOM vs DOC',
           xaxis = list(title = "DOC (mg/L)"),
           yaxis = list(title = "mean fDOM")
    )

```

## USDA-ARS Data - Little River

### Loading Data
Now let's start loading in the USDA-ARS Little River data that we downloaded earlier in the workshop. The files need to be in
your working directory. If you haven't already set this, you can do that with `setwd('path to directory')` in your script or console.
You can also set it by using the Files tab in RStudio. Browse to the directory you wish to use and then click the gear icon and 
click `Set As Working Directory`.

```{r lr data, echo = -1}
options(width = 100) #makes output wider in HTML
# Water chemistry data
lr.meta <- read.table('Metadata.tsv', 
                      header = TRUE, 
                      sep = '\t', 
                      fill = TRUE)
lr.all <- read.table('2078-B.txt', 
                 header = TRUE, 
                 sep = '\t')
head(lr.all)

```

You'll notice that the data contains a strange character (`·`) where
data is missing. We need to convert that to NA. We can copy it from the
data and use it as the `na.strings` value in `read.table`. There are a
varying number of spaces around that character so we'll also need to use
`strip.white` to remove the spaces.

```{r lr load data na}
lr.all <- read.table('2078-B.txt', 
                     header = TRUE, 
                     sep = '\t', 
                     strip.white = TRUE, 
                     na.strings = '·')
head(lr.all)

```

Now that we've figured that out, we can load in the discharge data.

```{r lr load meta q}
# Discharge data - just station B (6840), 2018
lr.q <- read.table('lrew_streamflow_daily.txt', 
                   header = TRUE, 
                   sep = ',', 
                   skip = 33, #contains lots of header rows, need to skip those
                   fill = TRUE)

```

### Process Data
The Litter River data includes data for many years and stations. Let's get just 2018 and station B (6840)

```{r lr select}
# Get columns of interest
lr <- lr.all[,c("STATION", "DATE", "TIME", "NO3N", "DOC")]

# Create proper date
lr$DATE <- as.Date(lr$DATE, format = '%m/%d/%Y')

# Only use 2018
lr <- lr[format(lr$DATE, '%Y') == '2018',]

# Discharge data - just station B (6840), 2018
lr.q <- lr.q[lr.q$ID == 6840,]
lr.q <- lr.q[lr.q$Year == 2018,]
```

There are a few last things we need to fix in the discharge data before we can use it. We need to change
the discharge date to a proper date format. Then we need to convert the discharge units to L/s to match NEON data.

```{r q units}
# Create date
lr.q$date <- as.Date(paste(lr.q$Year, lr.q$Month, lr.q$Day, sep = '-'))

# Convert Q cf/s to L/s to match NEON
lr.q$AvgDQlps <- lr.q$AvgDQ * 28.3168
```

### Plot Data
Now we're ready to start plotting.
```{r lr merge-plot}
# Plot nitrate, DOC over time. 
## We're saving the figure as a variable here so we can use it again below
fig <- plot_ly(data = lr, 
               x = ~DATE, 
               y = ~NO3N, 
               name = 'NO2+NO3-N', 
               type = 'scatter', 
               mode = 'lines+markers') %>%
          add_trace(x = ~DATE, 
                    y = ~DOC, 
                    name = 'DOC', 
                    type = 'scatter', 
                    mode = 'lines+markers') %>%
          layout(title = 'Little River: Nitrate and DOC',
                 xaxis = list(title = 'Date'),
                 yaxis = list(title = 'Concentration (mg/L)'))
fig #shows figure

# Plot all discharge
plot_ly(data = lr.q, 
        x = ~date, 
        y = ~AvgDQlps, 
        type = 'scatter', 
        mode = 'lines+markers') %>%
  layout(title = 'Little River: Discharge',
         xaxis = list(title = 'Date'),
         yaxis = list(title = 'Discharge (L/s)'))

# Plot all together
## need to make second y axis for discharge
y2 <- list(
  overlaying = 'y',
  side =  'right',
  title = 'Discharge (L/s)'
)

# Add discharge to fig from above
fig2 <- fig %>% 
  add_trace(data = lr.q, 
            x = ~date,
            y = ~AvgDQlps, 
            name = 'Discharge', 
            type = 'scatter', 
            mode = 'lines+markers', 
            yaxis = 'y2') %>%
  layout(title = 'Little River: Nitrate, DOC, Discharge',
         yaxis2 = y2)
fig2

```

## Comparison of Flint River and Little River
Let's plot FLNT and Little River data together
```{r plot both}
# Let's get daily averages of FLNT discharge so it matches Little River and so there isn't so much data to plot.
## First, we need to remove the time from endDate. We'll use floor_date from lubridate
csd_continuousDischarge$daily <- lubridate::floor_date(csd_continuousDischarge$endDate, unit = 'day')
## Next we will average by date
csd_daily <- setNames(aggregate(csd_continuousDischarge$maxpostDischarge, 
                                by = list(csd_continuousDischarge$daily), 
                                FUN = mean, 
                                na.rm = TRUE),
                      c('date','dailyQ'))

# Plot Discharge - multiply LR Q by 10 for easier comparison
plot_ly(data = csd_daily,
        x = ~date,
        y = ~dailyQ,
        type = 'scatter',
        mode = 'lines',
        name = 'FLNT') %>%
  add_trace(data = lr.q,
            x = ~date,
            y = ~AvgDQlps*10, #10x
            type = 'scatter',
            mode = 'lines',
            name = 'LR x 10') %>%
  layout(title = 'Discharge in both Rivers',
         xaxis = list(title = 'Date'),
         yaxis = list(title = 'Discharge (L/s)')
        )


# We've already processed the NEON water chem data as swc_nitrate and swc_doc. Let's just combine them together for plotting
swc_no3_doc <- cbind(swc_nitrate, swc_doc$analyteConcentration)
names(swc_no3_doc) <- c('collectDate','nitrate','roundedDate','DOC')

# Plot FLNT and LR Nitrate
plot_ly(data = swc_no3_doc, 
        x = ~collectDate, 
        y = ~nitrate, 
        type = 'scatter', 
        mode = 'lines+markers', 
        name = 'FLNT') %>%
  add_trace(data = lr, 
            x = ~DATE, 
            y = ~NO3N, 
            type = 'scatter', 
            mode = 'lines+markers', 
            name = 'LR') %>%
  layout(title = 'Nitrate in both Rivers',
         xaxis = list(title = 'Date'),
         yaxis = list(title = 'NO2+NO3-N (mg/L)'))

# Plot FLNT and LR DOC
plot_ly(data = swc_no3_doc, 
        x = ~collectDate, 
        y = ~DOC, 
        type = 'scatter', 
        mode = 'lines+markers', 
        name = 'FLNT') %>%
  add_trace(data = lr, 
            x = ~DATE, 
            y = ~DOC, 
            type = 'scatter', 
            mode = 'lines+markers', 
            name = 'LR') %>%
  layout(title = 'DOC in both Rivers',
         xaxis = list(title = 'Date'),
         yaxis = list(title = 'DOC (mg/L)'))

```

## Concentration-Discharge Plots For Both Sites
Finally, let's make C-Q plots for nitrate and DOC for both rivers. 
```{r qc plots}
# FLNT:
## Merge grab samples and discharge
mergedData.neon <- merge(csd_continuousDischarge,
                         swc_no3_doc,
                         by.x = "endDate",
                         by.y = "collectDate",
                         all.x = TRUE,
                         all.y = TRUE)

## Create regression of nitrate grab samples versus discharge and fit power function
concDisReg <- na.omit(mergedData.neon[,c('nitrate', 'DOC', 'maxpostDischarge')])
concDisReg <- concDisReg[order(concDisReg$maxpostDischarge),]
fit <- lm(log(nitrate) ~ log(maxpostDischarge), data=concDisReg)
plot_ly(data = concDisReg, 
        x = ~maxpostDischarge, 
        y = ~nitrate, 
        type = 'scatter',
        mode = 'markers', 
        name = 'data') %>%
  add_trace(data = concDisReg, 
            x = ~maxpostDischarge, 
            y = exp(predict(fit)),
            type = 'scatter',
            mode = 'lines', 
            name = 'regression') %>%
  layout(title = 'Flint River: Nitrate vs Q',
         xaxis = list(title = 'Q (L/s)'),
         yaxis = list(title = 'NO2+NO3-N (mg/L)'))

## Plot DOC vs Q
fit.doc <- lm(log(DOC) ~ log(maxpostDischarge), data = concDisReg)

plot_ly(data = concDisReg, 
        x = ~maxpostDischarge, 
        y = ~DOC, 
        type='scatter',
        mode = 'markers', 
        name = 'data') %>%
  add_trace(data = concDisReg, 
            x = ~maxpostDischarge, 
            y = exp(predict(fit.doc)),
            type = 'scatter',
            mode = 'lines',
            name = 'regression') %>%
  layout(title = 'Flint River: DOC vs Q',
         xaxis = list(title = 'Q (L/s)'),
         yaxis = list(title = 'DOC (mg/L)'))

# Little River:
## Merge discharge data and grab samples
merge.lr <- merge(lr, 
                  lr.q, 
                  by.x = 'DATE', 
                  by.y = 'date')

merge.lr <- merge.lr[order(merge.lr$AvgDQlps),]

## Plot nitrate vs Q
fit.lr.no3.q <- lm(log(NO3N) ~ log(AvgDQlps), data = merge.lr)

plot_ly(data = merge.lr, 
        x = ~AvgDQlps, 
        y = ~NO3N, 
        type = 'scatter', 
        mode = 'markers',
        name = 'data') %>%
  add_trace(data = merge.lr, 
            x = ~AvgDQlps, 
            y = exp(predict(fit.lr.no3.q)), 
            mode = 'lines',
            name = 'regression') %>%
  layout(title = 'Little River: Nitrate vs Q',
         xaxis = list(title = 'Q (L/s)'),
         yaxis = list(title = 'NO2+NO3-N (mg/L)'))


## Plot DOC vs Q
fit.lr.doc.q <- lm(log(DOC) ~ log(AvgDQlps), data = merge.lr)

plot_ly(data = merge.lr, 
        x = ~AvgDQlps, 
        y = ~DOC, 
        type = 'scatter', 
        mode = 'markers',
        name = 'data') %>%
  add_trace(data = merge.lr, 
            x = ~AvgDQlps, 
            y = exp(predict(fit.lr.doc.q)), 
            mode = 'lines',
            name = 'regression') %>%
    layout(title = 'Little River: DOC vs Q',
           xaxis = list(title = 'Q (L/s)'),
           yaxis = list(title = 'DOC (mg/L)'))

```

## Bonus - Get USGS data from R
You can use the `dataRetrieval` packages from the USGS to get access to [USGS Water Data](https://waterdata.usgs.gov). You'll need to know the site number and parameter code for the data you want to download. Below we pull in-situ nitrate and discharge data from the Santa Fe River in Florida. Then we merge it and plot the C-Q relationship.
```{r usgs, cache=TRUE}
# Load USGS dataRetrieval package. For more information visit: https://waterdata.usgs.gov/blog/dataretrieval/

#install.packages('dataRetrieval')
library(dataRetrieval)

# Pull discharge data (parameterCd=00060) and nitrate sensor data (parameterCd="99133")
# from the Santa Fe River (siteNumber=02322800). 
# note: the readNWISdv function returns daily values (dv). YYou could also use readNWISuv for instantaneous values

# Get daily values (dv) of discharge.
usgsDischarge<-dataRetrieval::readNWISdv(siteNumber="02322800", 
                                         parameterCd="00060", 
                                         startDate="2012-09-27",
                                         endDate="2016-09-30")

# Get daily values of nitrate sensor data 
usgsNitrate<-dataRetrieval::readNWISdv(siteNumber="02322800", 
                                       parameterCd="99133", 
                                       startDate="2012-09-27",
                                       endDate="2016-09-30")

# Merge data tables using dateTime column
usgsNitrate<-usgsNitrate[,c("Date","X_99133_00003","X_99133_00003_cd")]
usgsMerged<-merge(usgsDischarge,
                  usgsNitrate,
                  by = "Date")

# Plot C-Q relationship
plot_ly(data = usgsMerged, 
        x = ~X_00060_00003, 
        y = ~X_99133_00003, 
        type = 'scatter',
        mode = 'markers', 
        name = 'data') %>%
  layout(title = 'USGS: Santa Fe River In-site Nitrate vs Q',
         xaxis = list(title = 'Q (cfs)'),
         yaxis = list(title = 'NO2+NO3-N (mg/L)'))
```